YouTube Kids: The Effect of YouTube on Children’s Racial Learning

### 2025-03-10 manuscript.qmd
updated manuscript draft

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
floatsintext: true
numbered-lines: false
bibliography: [r-references.bib, bibliography.bib]
link-citations: true
draft-date: false
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
format:
  apaquarto-html: default
  apaquarto-docx: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    list-of-tables: true
    fig-width: 2
    fig-height: 2
    out-width: 40%
editor: visual
---

@@

library(svglite)
library(papaja)
library(knitr)
library(broom)
library(kableExtra)
library(psych)
library(tidyverse)

@@

@@

### Participants and Procedures {#sec-participants-and-procedures}
This study includes dyads of parents and children between the ages of 8 and 13 [^1]. Though children may express negative racial attitudes prior to this point in development, during early adolescence, children develop the ability to recognize and reflect on the racial patterns and dynamics present in their environments [@hugheswatford_2022_racial]. As such, this age range allows me to observe how the observation racial regularities present on YouTube may affect children’s stereotyping and prejudicial attitudes, and how such a phenomenon develops. Participants were recruited through Qualtrics Panels, allowing me to obtain a representative sample of parents and children; other studies have used Qualtrics Panels with similar methods and populations [@byrdahn_2020_profiles].

@@

Pilot data has been collected in March of 2025 from families recruited via Qualtrics Panels (*N* = `r length(transaction_id)`). Participating families receive two Qualtrics surveys to be completed asynchronously, and receive compensation (as determined by Qualtrics Panels) after both surveys have been completed. First, parents complete a demographics and YouTube usage survey, then children complete their own survey.

@@

@@

alpha(youtube %>% select(malleable_1:malleable_4))
alpha(youtube %>% select(homophily_appear_1:homophily_appear_5))
alpha(youtube %>% select(homophily_behave_1:homophily_behave_6))
alpha(youtube %>% select(parasocial_1:parasocial_5))
alpha(youtube %>% select(regularities_1:regularities_5))

return_alpha <- function(data, cols) {
  # select data 
  selected_data <- data %>% select({{ cols }})
  # compute alpha
  alpha_result <- psych::alpha(selected_data, check.keys = TRUE)
  # Extract and return the raw alpha value
  return(alpha_result$total$raw_alpha)
}

reliability <- 
  return_alpha(malleable_1:malleable_4) 
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "Reliability of each scale used"
#| echo: false
#| message: false
#| warning: false

# Generate kable table
kable(reliability, format = "html", caption = "Function Outputs in One Column") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))

@@

@@

@@

@@

```{r}
#| label: regression-plot
#| fig-cap: "Simple Correlation Between Time Spent on YouTube and Racial Attitudes score"
#| fig-scap: "Time v. Attitudes"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4
#| echo: false
#| message: false
#| warning: false

model_plot <- ggplot(youtube) +
  aes(x = Ptime, y = malleable_score) +
  geom_point() +
  stat_smooth(method = "lm") +
  scale_y_continuous(name = "Racial Attitudes Score") +
  scale_x_continuous(name = "Time spent on YouTube")
model_plot

>>>>

---
title: "YouTube Kids: The Effect of YouTube on Children’s Racial Learning"
shorttitle: "YouTube Racial Attitudes"
author: 
- name: Reilly Amera
  corresponding: true
  email: ramera@uchicago.edu
  affiliations:
  - id: id1
    name: "University of Chicago"
    department: MA Program in the Social Sciences
floatsintext: true
numbered-lines: false
bibliography: [r-references.bib, MA_THESIS.bib]
link-citations: true
draft-date: false
lang: en
language:
  citation-last-author-separator: "and"
  citation-masked-author: "Masked Citation"
  citation-masked-date: "n.d."
  citation-masked-title: "Masked Title"
  email: "Email"
format:
  apaquarto-html: default
  apaquarto-docx: default
  apaquarto-typst: 
    keep-typ: true
    list-of-figures: true
    list-of-tables: true
    toc: true
  apaquarto-pdf:
    documentmode: man
    keep-tex: true
    list-of-figures: true
    list-of-tables: true
    fig-width: 2
    fig-height: 2
    out-width: 40%
editor: visual
---

@@

library(svglite)
library(papaja)
library(knitr)
library(broom)
library(kableExtra)
library(flextable)
library(psych)
library(tidyverse)

@@

```{r}
#| label: my-theme
#| echo: false
#| message: false
#| warning: false

my_theme <- function(){
  theme(
    # Plot background color 
    plot.background = element_rect(fill = NA),
    # Adding border 
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    # Panel background
    panel.background = element_rect(fill = NA),
    # Plot labels
    plot.title = element_text(color = "black", family = "serif", face = "bold"),
    plot.subtitle = element_text(color = "black", family = "serif"),
    plot.caption = element_text(color = "black", family = "serif", face = "italic"),
    # Customizing grid lines 
    panel.grid.major.y = element_line(color = "lightgrey", linetype = 1, linewidth = 0.5),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(color = "lightgrey", linetype = 1, linewidth = 0.5),
    panel.grid.minor.x = element_blank(),
    # Customizing facets
    strip.background = element_rect(color = "black", fill = NA),
    strip.text = element_text(color = "black", hjust = 0.5, family = "serif"),
    # Customizing axes
    axis.text.y = element_text(color = "black", family = "serif", face = "bold.italic"),
    axis.text.x = element_text(color = "black", family = "serif", face = "bold.italic"),
    axis.title = element_text(color = "black", family = "serif", face = "bold"),
    axis.ticks = element_line(color = "black"),
    # Customizing legend
    legend.title = element_text(color = "black", family = "serif", face = "italic"), 
    legend.text = element_text(color = "black", family = "serif")
  )
}

@@

### Participants and Procedures {#sec-participants-and-procedures}

This study includes dyads of parents and children between the ages of 8 and 13 [^1]. Though children may express negative racial attitudes prior to this point in development, during early adolescence, children develop the ability to recognize and reflect on the racial patterns and dynamics present in their environments [@hugheswatford_2022_racial]. As such, this age range allows me to observe how the observation racial regularities present on YouTube may affect children’s stereotyping and prejudicial attitudes, and how such a phenomenon develops. Participants were recruited through Qualtrics Panels, allowing me to obtain a representative sample of parents and children; other studies have used Qualtrics Panels with similar methods and populations [@byrdahn_2020_profiles].

@@

Pilot data has been collected in March of 2025 from families recruited via Qualtrics Panels (*N* = `r length(youtube$transaction_id)`). Participating families receive two Qualtrics surveys to be completed asynchronously, and receive compensation (as determined by Qualtrics Panels) after both surveys have been completed. First, parents complete a demographics and YouTube usage survey, then children complete their own survey.

@@

The particular measures used in the child survey are as follows:

1. Beliefs about groups (adapted from @halperinetal_2012_promoting)
2. Perceived homophily in behavior and appearance (adapted from @mccroskeyetal_1975_development)
3. Parasocial interaction with a YouTuber (adapted from @sungetal_2023_shes)
4. Racial regularities (created based on @rollinsetal_2022_who)
5. Friend choice (adapted from @amemiyabian_2024_why)
All measures used in adult studies have been specifically adapted for child readability.

@@

return_alpha <- function(data, cols) {
  # select data 
  selected_data <- data %>% select(all_of(cols))
  # compute alpha
  alpha_result <- psych::alpha(selected_data, check.keys = TRUE)
  # Extract and return the raw alpha value
  return(alpha_result$total$raw_alpha)
}
```

```{r}
#| label: reliability-df
#| echo: false
#| message: false
#| warning: false

measures <-
  list(
  "Group Malleability" = c("malleable_1", "malleable_2", "malleable_3", "malleable_4"),
  "Homophily (Appearance)" = c("homophily_appear_1", "homophily_appear_2", "homophily_appear_3", "homophily_appear_4", "homophily_appear_5"),
  "Homophily (Appearance)" = c("homophily_behave_1", "homophily_behave_2", "homophily_behave_3", "homophily_behave_4", "homophily_behave_6"),
  "Parasociality" = c("parasocial_1", "parasocial_2", "parasocial_3", "parasocial_4", "parasocial_5"),
  "Racial Regularities" = c("regularities_1", "regularities_2", "regularities_3", "regularities_4", "regularities_5")
)

reliability <- data.frame(
  Measure = names(measures),
  Alpha = sapply(measures, function(cols) return_alpha(youtube, cols))
)

reliability <- reliability %>%
  mutate(Alpha = round(Alpha, digits = 2)) %>% 
  mutate(
    Evaluation = case_when(
      Alpha >= 0.95 ~ "Excellent",
      Alpha >= 0.9 ~ "Great",
      Alpha >= 0.8 ~ "Good",
      Alpha >= 0.7 ~ "Acceptable",
      Alpha >= 0.6 ~ "Questionable",
      Alpha < 0.6 ~ "Poor"
    )
  )
```

```{r}
#| label: tbl-reliability
#| tbl-cap: "Questionnaire Reliability"
#| echo: false
#| message: false
#| warning: false

# Generate table
kable(reliability, 
      align = c("c", "c", "c"),
      caption = "Reliability (assessed through Cronbach's Alpha) of questionnaires used, within sample") %>% 
  kable_styling(position = "center", latex_options = c("striped"))

@@

@tbl-reliability reports the reliability of all questionnaires used, within the present sample.

@@

Looking first to the distributions of time spent on YouTube among  participants, just as many children use YouTube every day as (*N* = `r sum(youtube$daily == 'Every Day')`) as those who do not (*N* = `r sum(youtube$daily == 'Not Every Day')`). And, as shown in @fig-time-distribution, children who use YouTube every day appear to use YouTube for more time on an average day than children who do not. On average, children spend about `r round(mean(youtube$Ptime), digits = 0)` minutes, or `r round(mean(youtube$Ptime)/60, digits = 0)` hour, on YouTube on an average day.

```{r}
#| label: fig-time-distribution
#| fig-cap: "Distribution of watch time for kids"
#| fig-align: "center"
#| echo: false
#| message: false
#| warning: false

# Set figure output settings - renders really small

ggplot(subset(youtube, !is.na(Ctime)), aes(x = Ctime)) + 
  geom_bar(color = "black", fill = "grey") +
  facet_grid(~ daily) +
  scale_x_discrete(labels = c("About 15 minutes a day" = "~ 15 min", "About 30 minutes a day" = "~ 30 min", "About 45 minutes a day" = "~ 45 min", "About 1 hour a day" = "~ 1 hr", "About 1 and 1/2 hours a day" = "~ 1.5 hr", "More than 2 hours a day" = "> 2 hr")) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(x = "Time Spent on YouTube",
       y = "Number of Participants",
       title = "Time Spent on YouTube on an Average Day",
       subtitle = "Comparing children who use YouTube every day and children who do not") +
  my_theme()

To what extent is time spent on YouTube predictive of how often children observe racial regularities?

```{r}
summary(aov(regularity_score ~ Ctime, data = youtube))
summary(aov(regularity_score ~ outgroup_freq, data = youtube))
```

```{r}
summary(aov(malleable_score ~ Ctime, data = youtube))
summary(aov(malleable_score ~ outgroup_freq, data = youtube))
```

To what extent is time spent on YouTube predictive

@@

summary(lm(malleable_score ~ regularity_score, data = youtube))

t.test(malleable_score ~ race, data = youtube)
t.test(regularity_score ~ race, data = youtube)

t.test(malleable_score ~ daily, data = youtube)
t.test(regularity_score ~ daily, data = youtube)
```

@@

```{r}
#| label: fig-regression-plot
#| fig-cap: "Simple Correlation Between Time Spent on YouTube and Racial Attitudes score"
#| fig-scap: "Time v. Attitudes"
#| fig-align: "center"
#| fig-width: 6
#| fig-height: 4
#| echo: false
#| message: false
#| warning: false

ggplot(youtube) +
  aes(x = regularity_score, y = malleable_score) +
  geom_point(aes(color = daily)) +
  stat_smooth(method = "lm") +
  facet_wrap(~ race) +
  scale_y_continuous(name = "Malleability Score") +
  scale_x_continuous(name = "Regularities Score") +
  my_theme()
